Cover Times of Random Walks

### 2019-04-09 paper.tex
matthews method

If we let the label of each graph in \cref{fig:example}
denote the expected cover time from that graph, we have the following system
of equations:
\begin{align}
\cov &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b} \nonumber \\
\textrm{a} &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
\textrm{b} &= 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e} \nonumber \\
\textrm{c} &= 1 + 1\textrm{e} \nonumber\\
\textrm{d} &= 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f} \nonumber \\
\textrm{e} &= 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
\textrm{f} &= 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0\nonumber \\
\textrm{g} &= 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber \\
\end{align}

@@

\subsection{Complete Graph}
We call the graph with an edge between each pair of vertices the
complete graph.
Note that a complete graph with $n$ vertices has ${n \choose 2}$ edges.

@@

\section{Bounds}
\subsection{Matthews Method}

$\hit \leq \cov$

@@

\begin{proof}
Without loss of generality, we may assume that the walk started at vertex $n$.
Let $\sigma$ be a uniform permutation of the unvisited $n-1$ vertices.
The strategy is to look for the vertices in order $\sigma$.
We call $\sigma_i$ the $i^{th}$ vertex we visit.
Let $T_k$ be the first time that vertices $\sigma_1, \sigma_2, ..., \sigma_k$
have all been visited.
For $1 \leq s \leq n-1$,
\begin{align}
\E_n[T_1 | \sigma_1 = s] = t_{n,s} \leq \hit \nonumber
\end{align}
\end{proof}

\subsection{Spanning Tree Argument}

\section{Distributional Aspects}
\subsection{Complete Graph}

\section{Electrical Networks}
\subsection{Definitions}
\subsection{Commute Times}
\subsection{Balanced Trees}

@@

\cite{AF14}

\cite{Bo98}
\cite{DS84}
\cite{Du11}
\cite{Gr10}
\cite{LP17}

>>>>

If we let the label of each graph in \cref{fig:example}
denote the expected cover time from that graph, we have the following system
of equations:
\begin{align}
\cov &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b} \nonumber \\
\textrm{a} &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
\textrm{b} &= 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e} \nonumber \\
\textrm{c} &= 1 + 1\textrm{e} \nonumber\\
\textrm{d} &= 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f} \nonumber \\
\textrm{e} &= 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
\textrm{f} &= 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0\nonumber \\
\textrm{g} &= 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber
\end{align}

@@

\subsection{Complete Graph \label{sec:complete}}
We call the graph with an edge between each pair of vertices the
complete graph.
Note that a complete graph with $n$ vertices has ${n \choose 2}$ edges.

@@

\section{Bounds}
\subsection{Matthews Method}
We have seen that it is in general very hard
to calculate the expected cover time of a random walk
on a graph.
The exceptions are symmetric or small graphs.
In lieu of an exact solution, we want to 
bound the expected cover time.

The Matthews Method is an example of a remarkably good bound.
In fact, it is tight for the complete graph:
the expected time between any two vertices
is $n-1$ so $\hit = n-1$ and \cref{thm:matthews} gives
us the main result of \cref{sec:complete}.

In addition, the proof uses strategies we employed on the
simple cycle and follows an intuitive structure.

@@

\begin{proof}
Without loss of generality, label the vertices $\{1, 2, ..., n\}$.
We may assume that the walk started at vertex $n$.
Let $\sigma$ be a uniform permutation of the unvisited $n-1$ vertices.
The strategy is to look for the vertices in order $\sigma$.

Let $T_k$ be the first time that vertices $\sigma_1, \sigma_2, ..., \sigma_k$
have all been visited.
For ease of notation, set $L_k = X_{T_k}$ to be the last state among 
$\sigma_1, \sigma_2, ..., \sigma_k$ to be visited (the first
time they are all visited).

We begin by writing the expected cover time in a clever way
(similar to the strategy we employed on the cycle graph):
\begin{align}
\cov &= \E_n[{T_{n-1}}]  \nonumber \\
&= \E_n[T_1 + (T_2 - T_1) + ... + 
(T_k - T_{k-1}) + ... + (T_{n-1} - T_{n-2})] . \nonumber
\end{align}

Our goal is to bound the cover time in terms of the hitting time.

Consider the first term of the sum.
Recall that the hitting time $\hit$ is the maximum
expected time from any vertex to any other.
It follows that for any choice of $\sigma_1$,
the expected time from $n$ to $\sigma_1$ is less than or equal to $\hit$.
Formally for $s$ an arbitrary state such that $1\leq s \leq n-1$,
$\E_n[T_1] = \E_n[T_s] \leq \hit$.

Now consider the $k^{th}$ term of the sum for $1 < k \leq n-1$.
By conditional expectation,

\begin{align}
\E_n[T_k - T_{k-1}] &=
\sum_{i=1}^k \E_n[T_k - T_{k-1} | L(k) = \sigma_i]
P(L_k = \sigma_i). \nonumber 
\end{align}

Assume that the last vertex $L_k$ to be visted among the first
$k$ vertices is not $\sigma_k$.
Then the first time we visit the first $k-1$ vertices
is also the first time we visit the first $k$ vertices
(since we must have already visited $\sigma_k$).
Formally, $\E_n[T_k-T_{k-1}|L(k) \neq \sigma_k] = 0$.
Then

\begin{align}
\E_n[T_k - T_{k-1}] &= 0 +
\E_n[T_k - T_{k-1} | L(k) = \sigma_k]
P(L_k = \sigma_k). \nonumber 
\end{align}

Since $\sigma$ is a random permutation, the probability that
we visit $\sigma_k$ last is $\frac{1}{k}$.
Formally, $P(L_k = \sigma_k) = \frac{1}{k}$.

Finally, consider the expected time between $T_{k-1}$ and $T_{k}$
given that last vertex we visit is $\sigma_k$.
Certainly there exist $r$ and $s$ where $1 \leq r \neq s \leq n-1$
such that $L_{k-1} = r$ and $L_k = s$.
Then $\E_n[T_k - T_{k-1} | L(k) = \sigma_k] = \E_r[T_s] \leq \hit$.
So
\begin{align}
\E_n[T_k - T_{k-1}] &\leq
\hit \frac{1}{k}. \nonumber 
\end{align}

Putting it all together,
\begin{align}
\cov \leq \hit \left( 1+ \frac{1}{2} + ... + \frac{1}{k} 
+ ... + \frac{1}{n-1} \right). \nonumber
\end{align}

\end{proof}

\subsection{Spanning Tree Argument}
\cite{AF14} 3.2

Edge commute inequality

spanning trees

\section{Distributional Aspects}
\subsection{Complete Graph}
\cite{Du11}
Theorem 2.2.4 -> Example 2.2.5 -> Theorem 3.6.5 -> Example 3.6.6

\section{Electrical Networks}
\subsection{Definitions}
\cite{AF14} 3.3
\subsection{Commute Time}
\cite{AF14} Proposition 2.3 -> Corollary 2.8 -> Corollary 3.11

\url{https://ocw.mit.edu/courses/mathematics/18-445-introduction-to-stochastic-processes-spring-2015/lecture-notes/MIT18_445S15_lecture8.pdf}

\subsection{Balanced Trees}
\cite{LP17} 11.3

@@

\cite{AF14}
\cite{Bo98}
\cite{DS84}
\cite{Du11}
\cite{Gr10}
\cite{LP17}
